---
title: "WPF Housing Instability"
author: "Anna Duan"
date: "`r Sys.Date()`"
output:
  html_document:
    keep_md: yes
    toc: yes
    toc_float: yes
    code_folding: hide
    number_sections: yes
    theme: "flatly"
  pdf_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE
)
library(tidyverse)
library(pander)
library(conflicted)
conflicts_prefer(dplyr::filter)
```

# Surveys  
Baseline survey: n = 5200
Survey: n = 6251
```{r survey universe}
survey_1 <- read.csv("/Users/annaduan/Library/CloudStorage/Box-Box/CLS Renter Survey/data/survey/Philly+Baseline+-+Phase+4_August+6,+2021_10.57.csv") %>%
  unique()

q1 <- survey_1 %>%
  head(1) %>%
  gather(key = "code", value = "question")

survey_2 <- read.csv("/Users/annaduan/Library/CloudStorage/Box-Box/CLS Renter Survey/data/survey/CLS+Renter+Survey_August+25,+2021_12.26.csv") %>%
  unique()

q2 <- survey_2 %>%
  head(1) %>%
  gather(key = "code", value = "question")

t <- full_join(q1, q2, by = "question")

```

# PUMS
PUMS: 38456 households, 117328 people
```{r pums universe}
pums <- read.csv("data/pums2021.csv")
```

### Philadelphia overall
#### Overall
##### Housing cost burden
```{r housing cost burden}
pums %>%
    mutate(count = 1,
      rent_burden_cat = factor(case_when(
      GRPIP < 30 ~ "Low (< 30%)",
      GRPIP >= 30 & GRPIP < 50 ~ "Moderate (30-50%)",
      GRPIP >= 50 ~ "Severe (>= 50%)"
    ))) %>%
  group_by(rent_burden_cat) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(pct)) %>%
  pander(caption = "Rent burden")
```

##### Overcrowding 
```{r overcrowding}
pums %>%
  select(NP, BDSP) %>%
  mutate(count = 1,
      occupants_per_room = factor(case_when(
      NP/BDSP <= 2 ~ "<=2",
      NP/BDSP >= 2 & NP/BDSP < 6 ~ "3-5",
      NP/BDSP >= 6 & NP/BDSP < 9 ~ "6-9",
      NP/BDSP >= 9 ~ "9+"
    ))) %>%
  group_by(occupants_per_room) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(pct)) %>%
  pander(caption = "Occupants per room")
```
##### Residence 1 year ago
```{r residence 1 year ago}
pums %>%
  select(MIG_label) %>%
  mutate(count = 1) %>%
  group_by(MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(pct)) %>%
  pander(caption = "Moved in last year")
```
##### Tenure
```{r tenure}
pums %>%
  select(TEN_label) %>%
  mutate(count = 1) %>%
  group_by(TEN_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(pct)) %>%
  pander(caption = "Tenure")
```
##### Housing quality
```{r housing quality}
pums %>%
  mutate(count = 1,
         housing_quality = factor(case_when(
      KIT == 1 & PLM == 1 ~ "full kitchen, full plumbing",
      KIT == 1 & PLM == 2 ~ "full kitchen, partial plumbing",
      KIT == 2 & PLM == 1 ~ "partial kitchen, full plumbing",
      KIT == 2 & PLM == 2 ~ "partial kitchen, partial plumbing"
    ))) %>%
  group_by(housing_quality) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(pct)) %>%
  pander(caption = "Housing quality")
```

#### Income 
##### Housing cost burden
```{r income/housing cost burden}
pums_inc <- pums %>%
  mutate(count = 1,
      income_cat = factor(case_when(
      HINCP < 35000 ~ "< 35k",
      HINCP >= 35000 & HINCP < 50000 ~ "35k - 50k",
      HINCP >= 50000 & HINCP < 100000 ~ "50k - 100k",
      HINCP >= 100000 ~ "> 100k"
    )))

pums_inc %>%
    mutate(rent_burden_cat = factor(case_when(
      GRPIP < 30 ~ "Low (< 30%)",
      GRPIP >= 30 & GRPIP < 50 ~ "Moderate (30-50%)",
      GRPIP >= 50 ~ "Severe (>= 50%)"
    ))) %>%
  group_by(income_cat, rent_burden_cat) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(income_cat) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(income_cat), desc(rent_burden_cat)) %>%
  pander(caption = "Rent burden")
```
##### Overcrowding
```{r income/overcrowding}
pums_inc %>%
  mutate(count = 1,
    occupants_per_room = factor(case_when(
      NP/BDSP <= 2 ~ "<=2",
      NP/BDSP >= 2 & NP/BDSP < 6 ~ "3-5",
      NP/BDSP >= 6 & NP/BDSP < 9 ~ "6-9",
      NP/BDSP >= 9 ~ "9+"
    ))) %>%
  group_by(income_cat, occupants_per_room) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(income_cat) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(income_cat), desc(occupants_per_room)) %>%
  pander(caption = "Occupants per room")
```
##### Residence 1 year ago
```{r income/residence 1 year ago}
pums_inc %>%
  group_by(income_cat, MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(income_cat) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(income_cat), desc(MIG_label)) %>%
  pander(caption = "Moved in last year")
```
##### Tenure
```{r income/tenure}
pums_inc %>%
  group_by(income_cat, TEN_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(income_cat) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(income_cat), desc(TEN_label)) %>%
  pander(caption = "Tenure")
```

##### Housing quality
```{r income/housing quality}
pums_inc %>%
    mutate(count = 1,
         housing_quality = factor(case_when(
      KIT == 1 & PLM == 1 ~ "complete facilities",
      KIT == 2 | PLM == 2 ~ "incomplete facilities"
    ))) %>%
  group_by(income_cat, housing_quality) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(income_cat) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(income_cat), desc(housing_quality)) %>%
  pander(caption = "Housing quality")
```

#### Race/ethnicity
##### Rent burden
```{r race/housing cost burden}
pums %>%
    mutate(rent_burden_cat = factor(case_when(
      GRPIP < 30 ~ "Low (< 30%)",
      GRPIP >= 30 & GRPIP < 50 ~ "Moderate (30-50%)",
      GRPIP >= 50 ~ "Severe (>= 50%)"
    )),
    count = 1) %>%
  group_by(HHLDRRAC1P_label, rent_burden_cat) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(HHLDRRAC1P_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(HHLDRRAC1P_label), desc(rent_burden_cat), desc(pct)) %>%
  pander(caption = "Rent burden")
```
##### Overcrowding
```{r race/overcrowding}
pums %>%
  mutate(count = 1,
    occupants_per_room = factor(case_when(
      NP/BDSP <= 2 ~ "<=2",
      NP/BDSP >= 2 & NP/BDSP < 6 ~ "3-5",
      NP/BDSP >= 6 & NP/BDSP < 9 ~ "6-9",
      NP/BDSP >= 9 ~ "9+"
    ))) %>%
  group_by(HHLDRRAC1P_label, occupants_per_room) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(HHLDRRAC1P_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(HHLDRRAC1P_label), desc(occupants_per_room)) %>%
  pander(caption = "Occupants per room")
```
##### Residence 1 year ago
```{r race/residence 1 year ago}
pums %>%
  mutate(count = 1) %>%
  group_by(HHLDRRAC1P_label, MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(HHLDRRAC1P_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(HHLDRRAC1P_label), desc(MIG_label)) %>%
  pander(caption = "Residence 1 year ago")
```
##### Tenure 
```{r race/tenure}
pums %>%
  mutate(count = 1) %>%
  group_by(HHLDRRAC1P_label, MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(HHLDRRAC1P_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(HHLDRRAC1P_label), desc(MIG_label)) %>%
  pander(caption = "Residence 1 year ago")
```
##### Housing quality
```{r race/housing quality}
pums %>%
    mutate(count = 1,
         housing_quality = factor(case_when(
      KIT == 1 & PLM == 1 ~ "full kitchen, full plumbing",
      KIT == 1 & PLM == 2 ~ "full kitchen, partial plumbing",
      KIT == 2 & PLM == 1 ~ "partial kitchen, full plumbing",
      KIT == 2 & PLM == 2 ~ "partial kitchen, partial plumbing"
    ))) %>%
  group_by(HHLDRRAC1P_label, housing_quality) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(HHLDRRAC1P_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(HHLDRRAC1P_label), desc(housing_quality)) %>%
  pander(caption = "Housing quality")
```

#### Sub-geography
#### Housing cost burden
```{r geo/housing cost burden}
pums_puma <- pums %>%
  mutate(region = factor(case_when(
    PUMA == 03201  ~ "Far Northeast",
    PUMA == 03202  ~ "Far Northeast - West",
    PUMA == 03203  ~ "Far Northeast - East",
    PUMA == 03204  ~ "North",
    PUMA == 03205  ~ "East",
    PUMA == 03206  ~ "Northwest",
    PUMA == 03207  ~ "Central",
    PUMA == 03208  ~ "West",
    PUMA == 03209  ~ "Center City",
    PUMA == 03210  ~ "Southwest",
    PUMA == 03211  ~ "Southeast")))

pums_puma %>%
   mutate(rent_burden_cat = factor(case_when(
      GRPIP < 30 ~ "Low (< 30%)",
      GRPIP >= 30 & GRPIP < 50 ~ "Moderate (30-50%)",
      GRPIP >= 50 ~ "Severe (>= 50%)"
    )),
    count = 1) %>%
  group_by(region, rent_burden_cat) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(region) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(region), desc(rent_burden_cat), desc(pct)) %>%
  pander(caption = "Rent burden")
```
##### Overcrowding
```{r geo/overcrowding}
pums_puma %>%
  mutate(count = 1,
    occupants_per_room = factor(case_when(
      NP/BDSP <= 2 ~ "<=2",
      NP/BDSP >= 2 & NP/BDSP < 6 ~ "3-5",
      NP/BDSP >= 6 & NP/BDSP < 9 ~ "6-9",
      NP/BDSP >= 9 ~ "9+"
    ))) %>%
  group_by(region, occupants_per_room) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(region) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(region), desc(occupants_per_room)) %>%
  pander(caption = "Occupants per room")
```

##### Residence 1 year ago
```{r geo/residence 1 year ago}
pums_puma %>%
  mutate(count = 1) %>%
  group_by(region, MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(region) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(region), desc(MIG_label)) %>%
  pander(caption = "Residence 1 year ago")
```

##### Tenure
```{r geo/tenure}
pums_puma %>%
  mutate(count = 1) %>%
  group_by(region, MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(region) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(region), desc(MIG_label)) %>%
  pander(caption = "Residence 1 year ago")
```

##### Housing quality
```{r geo/housing quality}
pums_puma %>%
    mutate(count = 1,
         housing_quality = factor(case_when(
      KIT == 1 & PLM == 1 ~ "complete facilities",
      KIT == 2 | PLM == 2 ~ "incomplete facilities"
    ))) %>%
  group_by(region, housing_quality) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(region) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(region), desc(housing_quality)) %>%
  pander(caption = "Housing quality")
```

#### Tenure
##### Housing cost burden
```{r tenure/housing cost burden}
pums %>%
   mutate(rent_burden_cat = factor(case_when(
      GRPIP < 30 ~ "Low (< 30%)",
      GRPIP >= 30 & GRPIP < 50 ~ "Moderate (30-50%)",
      GRPIP >= 50 ~ "Severe (>= 50%)"
    )),
    count = 1) %>%
  group_by(TEN_label, rent_burden_cat) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(TEN_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(TEN_label), desc(rent_burden_cat), desc(pct)) %>%
  pander(caption = "Rent burden")
```

##### Overcrowding
```{r tenure/overcrowding}
pums %>%
  mutate(count = 1,
    occupants_per_room = factor(case_when(
      NP/BDSP <= 2 ~ "<=2",
      NP/BDSP >= 2 & NP/BDSP < 6 ~ "3-5",
      NP/BDSP >= 6 & NP/BDSP < 9 ~ "6-9",
      NP/BDSP >= 9 ~ "9+"
    ))) %>%
  group_by(TEN_label, occupants_per_room) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(TEN_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(TEN_label), desc(occupants_per_room)) %>%
  pander(caption = "Occupants per room")
```

##### Residence 1 year ago
```{r tenure/residence 1 year ago}
pums %>%
  mutate(count = 1) %>%
  group_by(TEN_label, MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(TEN_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(TEN_label), desc(MIG_label)) %>%
  pander(caption = "Residence 1 year ago")
```

##### Housing quality
```{r tenure/housing quality}
pums %>%
    mutate(count = 1,
         housing_quality = factor(case_when(
      KIT == 1 & PLM == 1 ~ "complete facilities",
      KIT == 2 | PLM == 2 ~ "incomplete facilities"
    ))) %>%
  group_by(TEN_label, housing_quality) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(TEN_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(TEN_label), desc(housing_quality)) %>%
  pander(caption = "Housing quality")
```

### Philadelphia households with young children (<6 years)
Households with children < 6 years: n = 3528
##### Housing cost burden
```{r housing cost burden x kids}
pums_kids <- pums %>%
  filter(HUPAC_label == "With children under 6 years only")

pums_kids %>%
    mutate(count = 1,
      rent_burden_cat = factor(case_when(
      GRPIP < 30 ~ "Low (< 30%)",
      GRPIP >= 30 & GRPIP < 50 ~ "Moderate (30-50%)",
      GRPIP >= 50 ~ "Severe (>= 50%)"
    ))) %>%
  group_by(rent_burden_cat) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(pct)) %>%
  pander(caption = "Rent burden")
```

##### Overcrowding 
```{r overcrowding x kids}
pums_kids %>%
  select(NP, BDSP) %>%
  mutate(count = 1,
      occupants_per_room = factor(case_when(
      NP/BDSP <= 2 ~ "<=2",
      NP/BDSP >= 2 & NP/BDSP < 6 ~ "3-5",
      NP/BDSP >= 6 & NP/BDSP < 9 ~ "6-9",
      NP/BDSP >= 9 ~ "9+"
    ))) %>%
  group_by(occupants_per_room) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(pct)) %>%
  pander(caption = "Occupants per room")
```
##### Residence 1 year ago
```{r residence 1 year ago x kids}
pums_kids %>%
  select(MIG_label) %>%
  mutate(count = 1) %>%
  group_by(MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(pct)) %>%
  pander(caption = "Moved in last year")
```
##### Tenure
```{r tenure x kids}
pums_kids %>%
  select(TEN_label) %>%
  mutate(count = 1) %>%
  group_by(TEN_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(pct)) %>%
  pander(caption = "Tenure")
```
##### Housing quality
```{r housing quality x kids}
pums_kids %>%
  mutate(count = 1,
         housing_quality = factor(case_when(
      KIT == 1 & PLM == 1 ~ "complete facilities",
      KIT == 2 | PLM == 2 ~ "incomplete facilities"
    ))) %>%
  group_by(housing_quality) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(pct)) %>%
  pander(caption = "Housing quality")
```

#### Income 
##### Housing cost burden
```{r income/housing cost burden x kids}
pums_inc <- pums_kids %>%
  mutate(count = 1,
      income_cat = factor(case_when(
      HINCP < 35000 ~ "< 35k",
      HINCP >= 35000 & HINCP < 50000 ~ "35k - 50k",
      HINCP >= 50000 & HINCP < 100000 ~ "50k - 100k",
      HINCP >= 100000 ~ "> 100k"
    )))

pums_inc %>%
    mutate(rent_burden_cat = factor(case_when(
      GRPIP < 30 ~ "Low (< 30%)",
      GRPIP >= 30 & GRPIP < 50 ~ "Moderate (30-50%)",
      GRPIP >= 50 ~ "Severe (>= 50%)"
    ))) %>%
  group_by(income_cat, rent_burden_cat) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(income_cat) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(income_cat), desc(rent_burden_cat)) %>%
  pander(caption = "Rent burden")
```
##### Overcrowding
```{r income/overcrowding x kids}
pums_inc %>%
  mutate(count = 1,
    occupants_per_room = factor(case_when(
      NP/BDSP <= 2 ~ "<=2",
      NP/BDSP >= 2 & NP/BDSP < 6 ~ "3-5",
      NP/BDSP >= 6 & NP/BDSP < 9 ~ "6-9",
      NP/BDSP >= 9 ~ "9+"
    ))) %>%
  group_by(income_cat, occupants_per_room) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(income_cat) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(income_cat), desc(occupants_per_room)) %>%
  pander(caption = "Occupants per room")
```
##### Residence 1 year ago
```{r income/residence 1 year ago x kids}
pums_inc %>%
  group_by(income_cat, MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(income_cat) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(income_cat), desc(MIG_label)) %>%
  pander(caption = "Moved in last year")
```
##### Tenure
```{r income/tenure x kids}
pums_inc %>%
  group_by(income_cat, TEN_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(income_cat) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(income_cat), desc(TEN_label)) %>%
  pander(caption = "Tenure")
```

##### Housing quality
```{r income/housing quality x kids}
pums_inc %>%
    mutate(count = 1,
         housing_quality = factor(case_when(
      KIT == 1 & PLM == 1 ~ "complete facilities",
      KIT == 2 | PLM == 2 ~ "incomplete facilities"
    ))) %>%
  group_by(income_cat, housing_quality) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(income_cat) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(income_cat), desc(housing_quality)) %>%
  pander(caption = "Housing quality")
```

#### Race/ethnicity
##### Rent burden
```{r race/housing cost burden x kids}
pums_kids %>%
    mutate(rent_burden_cat = factor(case_when(
      GRPIP < 30 ~ "Low (< 30%)",
      GRPIP >= 30 & GRPIP < 50 ~ "Moderate (30-50%)",
      GRPIP >= 50 ~ "Severe (>= 50%)"
    )),
    count = 1) %>%
  group_by(HHLDRRAC1P_label, rent_burden_cat) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(HHLDRRAC1P_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(HHLDRRAC1P_label), desc(rent_burden_cat), desc(pct)) %>%
  pander(caption = "Rent burden")
```
##### Overcrowding
```{r race/overcrowding x kids}
pums_kids %>%
  mutate(count = 1,
    occupants_per_room = factor(case_when(
      NP/BDSP <= 2 ~ "<=2",
      NP/BDSP >= 2 & NP/BDSP < 6 ~ "3-5",
      NP/BDSP >= 6 & NP/BDSP < 9 ~ "6-9",
      NP/BDSP >= 9 ~ "9+"
    ))) %>%
  group_by(HHLDRRAC1P_label, occupants_per_room) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(HHLDRRAC1P_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(HHLDRRAC1P_label), desc(occupants_per_room)) %>%
  pander(caption = "Occupants per room")
```
##### Residence 1 year ago
```{r race/residence 1 year ago x kids}
pums_kids %>%
  mutate(count = 1) %>%
  group_by(HHLDRRAC1P_label, MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(HHLDRRAC1P_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(HHLDRRAC1P_label), desc(MIG_label)) %>%
  pander(caption = "Residence 1 year ago")
```
##### Tenure 
```{r race/tenure x kids}
pums_kids %>%
  mutate(count = 1) %>%
  group_by(HHLDRRAC1P_label, MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(HHLDRRAC1P_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(HHLDRRAC1P_label), desc(MIG_label)) %>%
  pander(caption = "Residence 1 year ago")
```
##### Housing quality
```{r race/housing quality x kids}
pums_kids %>%
    mutate(count = 1,
         housing_quality = factor(case_when(
      KIT == 1 & PLM == 1 ~ "complete facilities",
      KIT == 2 | PLM == 2 ~ "incomplete facilities"
    ))) %>%
  group_by(HHLDRRAC1P_label, housing_quality) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(HHLDRRAC1P_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(HHLDRRAC1P_label), desc(housing_quality)) %>%
  pander(caption = "Housing quality")
```

#### Sub-geography
#### Housing cost burden
```{r geo/housing cost burden x kids}
pums_puma <- pums_kids %>%
  mutate(region = factor(case_when(
    PUMA == 03201  ~ "Far Northeast",
    PUMA == 03202  ~ "Far Northeast - West",
    PUMA == 03203  ~ "Far Northeast - East",
    PUMA == 03204  ~ "North",
    PUMA == 03205  ~ "East",
    PUMA == 03206  ~ "Northwest",
    PUMA == 03207  ~ "Central",
    PUMA == 03208  ~ "West",
    PUMA == 03209  ~ "Center City",
    PUMA == 03210  ~ "Southwest",
    PUMA == 03211  ~ "Southeast")))

pums_puma %>%
   mutate(rent_burden_cat = factor(case_when(
      GRPIP < 30 ~ "Low (< 30%)",
      GRPIP >= 30 & GRPIP < 50 ~ "Moderate (30-50%)",
      GRPIP >= 50 ~ "Severe (>= 50%)"
    )),
    count = 1) %>%
  group_by(region, rent_burden_cat) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(region) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(region), desc(rent_burden_cat), desc(pct)) %>%
  pander(caption = "Rent burden")
```
##### Overcrowding
```{r geo/overcrowding x kids}
pums_puma %>%
  mutate(count = 1,
    occupants_per_room = factor(case_when(
      NP/BDSP <= 2 ~ "<=2",
      NP/BDSP >= 2 & NP/BDSP < 6 ~ "3-5",
      NP/BDSP >= 6 & NP/BDSP < 9 ~ "6-9",
      NP/BDSP >= 9 ~ "9+"
    ))) %>%
  group_by(region, occupants_per_room) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(region) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(region), desc(occupants_per_room)) %>%
  pander(caption = "Occupants per room")
```

##### Residence 1 year ago
```{r geo/residence 1 year ago x kids}
pums_puma %>%
  mutate(count = 1) %>%
  group_by(region, MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(region) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(region), desc(MIG_label)) %>%
  pander(caption = "Residence 1 year ago")
```

##### Tenure
```{r geo/tenure x kids}
pums_puma %>%
  mutate(count = 1) %>%
  group_by(region, MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(region) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(region), desc(MIG_label)) %>%
  pander(caption = "Residence 1 year ago")
```

##### Housing quality
```{r geo/housing quality x kids}
pums_puma %>%
    mutate(count = 1,
         housing_quality = factor(case_when(
      KIT == 1 & PLM == 1 ~ "complete facilities",
      KIT == 2 | PLM == 2 ~ "incomplete facilities"
    ))) %>%
  group_by(region, housing_quality) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(region) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(region), desc(housing_quality)) %>%
  pander(caption = "Housing quality")
```

#### Tenure
##### Housing cost burden
```{r tenure/housing cost burden x kids}
pums_kids %>%
   mutate(rent_burden_cat = factor(case_when(
      GRPIP < 30 ~ "Low (< 30%)",
      GRPIP >= 30 & GRPIP < 50 ~ "Moderate (30-50%)",
      GRPIP >= 50 ~ "Severe (>= 50%)"
    )),
    count = 1) %>%
  group_by(TEN_label, rent_burden_cat) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(TEN_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(TEN_label), desc(rent_burden_cat), desc(pct)) %>%
  pander(caption = "Rent burden")
```

##### Overcrowding
```{r tenure/overcrowding x kids}
pums_kids %>%
  mutate(count = 1,
    occupants_per_room = factor(case_when(
      NP/BDSP <= 2 ~ "<=2",
      NP/BDSP >= 2 & NP/BDSP < 6 ~ "3-5",
      NP/BDSP >= 6 & NP/BDSP < 9 ~ "6-9",
      NP/BDSP >= 9 ~ "9+"
    ))) %>%
  group_by(TEN_label, occupants_per_room) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(TEN_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(TEN_label), desc(occupants_per_room)) %>%
  pander(caption = "Occupants per room")
```

##### Residence 1 year ago
```{r tenure/residence 1 year ago x kids}
pums_kids %>%
  mutate(count = 1) %>%
  group_by(TEN_label, MIG_label) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(TEN_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(TEN_label), desc(MIG_label)) %>%
  pander(caption = "Residence 1 year ago")
```

##### Housing quality
```{r tenure/housing quality x kids}
pums_kids %>%
    mutate(count = 1,
         housing_quality = factor(case_when(
      KIT == 1 & PLM == 1 ~ "complete facilities",
      KIT == 2 | PLM == 2 ~ "incomplete facilities"
    ))) %>%
  group_by(TEN_label, housing_quality) %>%
  summarize(households = sum(count)) %>%
  ungroup() %>%
  group_by(TEN_label) %>%
  mutate(pct = 100*households/sum(households)) %>%
  arrange(desc(TEN_label), desc(housing_quality)) %>%
  pander(caption = "Housing quality")
```
